On $K_2$-analogue of Serre problem for $\SL_4$

### 2018-08-29 paper.tex
Minor fixes

Now suppose  $0 \leq d \leq m+1$. Clearly \eqref{lem:bmp1} and \eqref{eq:defmpn} (in the case $1\leq d\leq m$) or~\cref{lem:bmp1} (in the case $d=0,m+1$) imply that
\begin{equation} \label{eq:verify-bmp1'} 
[S_\alpha(\xi), x_{\alpha'}(\eta)] = S_{\alpha+\alpha'}(N_{\alpha,\alpha'}\xi \eta),\ \xi \in B_d, \eta \in B_{m+1-d}.
\end{equation}

We need to introduce additional notation.
For $a,b\leq m+1$ we denote by $\bot(a, b)$ (resp. $\angle(a,b)$) the set consisting of all relations $[S_\alpha(\xi), S_{\alpha'}(\eta)] = 1$
for which $\xi \in B_a$, $\eta \in B_b$ and $\alpha$ and $\alpha'$ are orthogonal (resp. form a sharp angle). 
We denote by $\bot_0(d, m+1)$ the subset of $\bot(d, m+1)$
consisting of those relations for which, moreover, $\eta = t^{m+1}$.

\begin{lemma} \label{claim1} For $d\leq m+1$ relations $\bot_0(d, 1)$ and $\angle(d, m)$ imply $\angle(d, m+1)$. \end{lemma}
\begin{proof}
Without loss of generality we may assume that $\alpha' = \alpha + \beta$
  for some $\rA_3$-triple $(\alpha, \beta, \gamma)$.
Write $\eta = bt^{m+1}$ for some $b\in A$.
\begin{align*} 
[S_\alpha(\xi), S_{\alpha+\beta}(bt^{m+1})] = [S_\alpha(\xi), [x_{\alpha+\beta+\gamma}(b't^m), x_{-\gamma}(t)]] & \text{ by~\eqref{item:i} and~\eqref{eq:defmpn}}\\
= [[S_\alpha(\xi), x_{\alpha+\beta+\gamma}(b't^m)], {}^{x_{\alpha+\beta+\gamma}(b't^m)}\!x_{-\gamma}(t)] & \text{ by~$\bot_0(d, 1)$}\\
= 1
 & \text{ by~$\angle(d, m)$. \qedhere} \end{align*} 
\end{proof}

\begin{lemma} \label{claim2} For $1\leq d\leq m+1$ relations $\angle(d, m+1)$ imply $\bot_0(d, m+1)$. \end{lemma}
\begin{proof}
As before, without loss of generality we may assume $\alpha' = \gamma$ for some $\mathsf{A}_3$-triple $(\alpha, \beta, \gamma)$.
 \begin{align*}
 {}^{S_\gamma(t^d)}S_\alpha(bt^{m+1}) = {}^{S_\gamma(t^d)}[S_{-\beta}(b't^{m+1}), x_{\alpha+\beta}(1)] & \text{ by~\cref{lem:bmp1}}\\
 = [S_{-\beta}(b't^{m+1}), {}^{S_\gamma(t^d)}x_{\alpha+\beta}(1)]                                     
 & \text{ by $\angle(d, m+1)$}\\
 = [[x_{-\beta-\gamma}(b''t^{m+1-d}), S_\gamma(t^d)], {}^{S_\gamma(t^d)}x_{\alpha+\beta}(1)]          
 & \text{ by~\eqref{eq:verify-bmp1'}}\\
 = [x_{-\beta-\gamma}(b''t^{m+1-d)}), [S_\gamma(t^d), x_{\alpha+\beta}(1)]]                           
 & \text{ by~\eqref{eq:H1iii} and $(b_{m+1-d})$}\\
 = S_{\alpha}(b'''t^{m+1})                                                                            
 & \text{ by~\eqref{eq:verify-bmp1'}.} \end{align*}
Usual identities for structure constants imply (cf.~\cite[p.~12]{Re75}):
 $b'''=N_{-\beta-\gamma, \alpha+\beta+\gamma} \cdot N_{\gamma, \alpha+\beta} \cdot N_{-\beta-\gamma, \gamma} \cdot N_{-\beta, \alpha+\beta} \cdot b = b,$ from which the claim follows.
\end{proof}

\begin{lemma} \label{claim3}
 For $d\leq m+1$ relations $\bot_0(d, m+1)$, $\angle(d, d)$ and $\angle(m+1, m+1)$ imply $\bot(d, m+1)$.
\end{lemma}
\begin{proof}
 \begin{align*} 
 S_{\alpha+\beta+\gamma}(abt^{m+1}) = {}^{S_{-\beta}(t^d)}\!S_{\alpha+\beta+\gamma}(abt^{m+1}) 
 & \text{ by $\bot_0(m+1, d)$}\\
 = {}^{S_{-\beta}(t^d)}\![x_{\alpha+\beta}(b't^{m+1-d}), S_\gamma(at^d)] 
 & \text{ by~\eqref{eq:verify-bmp1'}} \\
 = [S_{\alpha}(b't^{m+1}) x_{\alpha+\beta}(b't^{m+1-d}), S_\gamma(at^d)] 
 & \text{ by~\eqref{eq:verify-bmp1'} and $\angle(d, d)$ }\\
 = {}^{S_{\alpha}(b't^{m+1})}\!S_{\alpha+\beta+\gamma}(abt^{m+1}) [S_{\alpha}(b't^{m+1}), S_\gamma(at^d)] 
 & \text{ by~\eqref{eq:H1ii} and~\eqref{eq:verify-bmp1'}} \\
 = S_{\alpha+\beta+\gamma}(abt^{m+1}) [S_{\alpha}(b't^{m+1}), S_\gamma(at^d)] 
 & \text{ by $\angle(m+1, m+1)$. \qedhere}
\end{align*}
\end{proof}

\begin{lemma}
Elements $S_\alpha(\xi)$ satisfy relations $(b_{m+1})$.
 \end{lemma}
\begin{proof}
We need to show that $S_\alpha(\xi)$ satisfy $\angle(d, m+1)$ and $\perp(d, m+1)$ for $d\leq m+1$.
\begin{enumerate}
 \item Relations $\angle(d, m+1)$, $d\leq m$ follow from~\cref{claim1}. 
 \item Relations $\bot(d, m+1)$ with $d\leq 0$ can verified by direct computation:
 \begin{align*} 
 [x_\alpha(\xi), S_{\gamma}(bt^{m+1})] = [x_\alpha(\xi), [x_{\beta+\gamma}(b't^m), x_{-\beta}(t)]] & \text{ by~\eqref{item:i},\eqref{eq:defmpn}} \\
 = [[x_\alpha(\xi), x_{\beta+\gamma}(b't^m)], {}^{x_{\beta+\gamma}(b't^m)}\!x_{-\beta}(t)] &
 \text{ by \eqref{eq:H1iii} and $(b_1)$} \\
 = {}^{x_{\beta+\gamma}(b't^m)}\![x_{\alpha+\beta+\gamma}(b't^m\xi), x_{-\beta}(t)] &
 \text{ by $(b_m)$ and $(c_m)$} \\
 = 1 & \text{ by $(b_m)$.}
\end{align*}
 \item Relations $\bot_0(d, m+1)$ with $1\leq d\leq m$ follows from~\cref{claim2}.
 \item Relations $\angle(m+1, m+1)$ follow from~\cref{claim1}, consequently relations $\bot_0(m+1, m+1)$ follow from~\cref{claim2}.
 \item Relations $\bot(d, m+1)$, $0\leq d\leq m+1$ follow from~\cref{claim3} and the previous two assertions.
\end{enumerate}
\begin{comment}
 \begin{enumerate}
\item \label{case:1} First consider the case $d \leq 0$. There are two further subcases.
 \begin{enumerate}
  \item \label{case:1a} Case $\alpha \not \perp \alpha'$. 
  Without loss of generality we may assume that $\alpha' = \alpha + \beta$
  for some $\rA_3$-triple $(\alpha, \beta, \gamma)$.
  Using~\eqref{eq:H1iii} and~\eqref{eq:bm} we obtain
   \begin{multline} \nonumber
   [S_\alpha(\xi), S_{\alpha+\beta}(bt^{m+1})] = [x_\alpha(\xi), [x_{\alpha+\beta+\gamma}(b't^m), x_{-\gamma}(t)]] = \\   
  = [[x_\alpha(\xi), x_{\alpha+\beta+\gamma}(b't^m)], {}^{x_{\alpha+\beta+\gamma}(b't^m)}\!x_{-\gamma}(t)] = 1 \text{ for some $b'\in A$ }.
  \end{multline}
  What this really proves: $PERP(d, 1)$ and $NONPERP(d, m)$ imply $NONPERP(d, m+1)$.
  \item Case $\alpha \perp \alpha'$. 
  As before, we set $\alpha' = \gamma$ and use~\eqref{eq:H1iii},\eqref{eq:bm} and~\eqref{eq:cm}:
  \begin{multline} \nonumber
   [S_\alpha(\xi), S_{\gamma}(bt^{m+1})] = [S_\alpha(\xi), [x_{\beta+\gamma}(b't^m), x_{-\beta}(t)]] = \\   
  = [[S_\alpha(\xi), x_{\beta+\gamma}(b't^m)], {}^{x_{\beta+\gamma}(b't^m)}\!x_{-\beta}(t)] = {}^{x_{\beta+\gamma}(b't^m)}\![x_{\alpha+\beta+\gamma}(b't^m\xi), x_{-\beta}(t)] = 1.
  \end{multline}
 \end{enumerate}

>>>>

Now suppose  $0 \leq d \leq m+1$. Clearly \eqref{lem:bmp1} and \eqref{eq:defmpn} (in the case $1\leq d\leq m$) or~\cref{lem:bmp1} (in the case $d=0,m+1$) imply that
\begin{equation} \label{eq:verify-bmp1'} 
[S_\alpha(\xi), S_{\alpha'}(\eta)] = S_{\alpha+\alpha'}(N_{\alpha,\alpha'}\xi \eta),\ \xi \in B_d, \eta \in B_{m+1-d}.
\end{equation}

We need to introduce additional notation.
For $a,b\leq m+1$ we denote by $\bot(a, b)$ (resp. $\angle(a,b)$) the family of all relations $[S_\alpha(\xi), S_{\alpha'}(\eta)] = 1$
for which $\xi \in B_a$, $\eta \in B_b$ and $\alpha$ and $\alpha'$ are orthogonal (resp. form a sharp angle). 
We denote by $\bot_0(a, b)$ the subset of $\bot(a, b)$
consisting of those relations for which $\xi = t^{a} \in B_a$.

\begin{lemma} \label{claim1} In $G_m$ relations $\bot_0(1, d)$ and $\angle(d, m)$ imply $\angle(d, m+1)$ for $d\leq m+1$ \end{lemma}
\begin{proof}
Without loss of generality we may assume that $\alpha' = \alpha + \beta$
  for some $\rA_3$-triple $(\alpha, \beta, \gamma)$.
Write $\eta = bt^{m+1}$ for some $b\in A$.
\begin{align*} 
[S_\alpha(\xi), S_{\alpha+\beta}(bt^{m+1})] = [S_\alpha(\xi), [x_{\alpha+\beta+\gamma}(b't^m), x_{-\gamma}(t)]] & \text{ by~\eqref{item:i} and~\eqref{eq:defmpn}}\\
= [[S_\alpha(\xi), x_{\alpha+\beta+\gamma}(b't^m)], {}^{x_{\alpha+\beta+\gamma}(b't^m)}\!x_{-\gamma}(t)] & \text{ by~$\bot_0(1, d)$}\\
= 1
 & \text{ by~$\angle(d, m)$. \qedhere} \end{align*} 
\end{proof}

\begin{lemma} \label{claim2} In $G_m$ relations $\angle(d, m+1)$ imply $\bot_0(d, m+1)$ for  $1\leq d\leq m+1$. \end{lemma}
\begin{proof}
As before, without loss of generality we may assume $\alpha' = \gamma$ for some $\mathsf{A}_3$-triple $(\alpha, \beta, \gamma)$.
 \begin{align*}
 {}^{S_\gamma(t^d)}S_\alpha(bt^{m+1}) = {}^{S_\gamma(t^d)}[S_{-\beta}(b't^{m+1}), x_{\alpha+\beta}(1)] & \text{ by~\cref{lem:bmp1}}\\
 = [S_{-\beta}(b't^{m+1}), {}^{S_\gamma(t^d)}x_{\alpha+\beta}(1)]                                     
 & \text{ by $\angle(d, m+1)$}\\
 = [[x_{-\beta-\gamma}(b''t^{m+1-d}), S_\gamma(t^d)], {}^{S_\gamma(t^d)}x_{\alpha+\beta}(1)]          
 & \text{ by~\eqref{eq:verify-bmp1'}}\\
 = [x_{-\beta-\gamma}(b''t^{m+1-d)}), [S_\gamma(t^d), x_{\alpha+\beta}(1)]]                           
 & \text{ by~\eqref{eq:H1iii} and $(b_{m+1-d})$}\\
 = S_{\alpha}(b'''t^{m+1})                                                                            
 & \text{ by~\eqref{eq:verify-bmp1'}.} \end{align*}
Usual identities for structure constants imply (cf.~\cite[p.~12]{Re75}):
 $b'''=N_{-\beta-\gamma, \alpha+\beta+\gamma} \cdot N_{\gamma, \alpha+\beta} \cdot N_{-\beta-\gamma, \gamma} \cdot N_{-\beta, \alpha+\beta} \cdot b = b,$ from which the claim follows.
\end{proof}

\begin{lemma} \label{claim3}
 In $G_m$ relations $\bot_0(d, m+1)$, $\angle(d, d)$ and $\angle(m+1, m+1)$ imply $\bot(d, m+1)$ for $d\leq m+1$.
\end{lemma}
\begin{proof}
 \begin{align*} 
 S_{\alpha+\beta+\gamma}(abt^{m+1}) = {}^{S_{-\beta}(t^d)}\!S_{\alpha+\beta+\gamma}(abt^{m+1}) 
 & \text{ by $\bot_0(d, m+1)$}\\
 = {}^{S_{-\beta}(t^d)}\![x_{\alpha+\beta}(b't^{m+1-d}), S_\gamma(at^d)] 
 & \text{ by~\eqref{eq:verify-bmp1'}} \\
 = [S_{\alpha}(b't^{m+1}) x_{\alpha+\beta}(b't^{m+1-d}), S_\gamma(at^d)] 
 & \text{ by~\eqref{eq:verify-bmp1'} and $\angle(d, d)$ }\\
 = {}^{S_{\alpha}(b't^{m+1})}\!S_{\alpha+\beta+\gamma}(abt^{m+1}) [S_{\alpha}(b't^{m+1}), S_\gamma(at^d)] 
 & \text{ by~\eqref{eq:H1ii} and~\eqref{eq:verify-bmp1'}} \\
 = S_{\alpha+\beta+\gamma}(abt^{m+1}) [S_{\alpha}(b't^{m+1}), S_\gamma(at^d)] 
 & \text{ by $\angle(m+1, m+1)$. \qedhere}
\end{align*}
\end{proof}

\begin{lemma}
Elements $S_\alpha(\xi)$ satisfy relations $(b_{m+1})$.
 \end{lemma}
\begin{proof}
We need to show that $S_\alpha(\xi)$ satisfy $\angle(d, m+1)$ and $\bot(d, m+1)$ for $d\leq m+1$.
\begin{enumerate}
 \item Relations $\angle(d, m+1)$, $d\leq m$ follow from~\cref{claim1}. 
 \item Relations $\bot(d, m+1)$ with $d\leq 0$ can verified by direct computation:
 \begin{align*} 
 [x_\alpha(\xi), S_{\gamma}(bt^{m+1})] = [x_\alpha(\xi), [x_{\beta+\gamma}(b't^m), x_{-\beta}(t)]] & \text{ by~\eqref{item:i},\eqref{eq:defmpn}} \\
 = [[x_\alpha(\xi), x_{\beta+\gamma}(b't^m)], {}^{x_{\beta+\gamma}(b't^m)}\!x_{-\beta}(t)] &
 \text{ by \eqref{eq:H1iii} and $(b_1)$} \\
 = {}^{x_{\beta+\gamma}(b't^m)}\![x_{\alpha+\beta+\gamma}(b't^m\xi), x_{-\beta}(t)] &
 \text{ by $(b_m)$ and $(c_m)$} \\
 = 1 & \text{ by $(b_m)$.}
\end{align*}
 \item Relations $\bot_0(d, m+1)$ with $1\leq d\leq m$ follow from~\cref{claim2}.
 \item Relations $\angle(m+1, m+1)$ follow from~\cref{claim1}, consequently relations $\bot_0(m+1, m+1)$ follow from~\cref{claim2}.
 \item Relations $\bot(d, m+1)$, $0\leq d\leq m+1$ follow from~\cref{claim3} and the previous two assertions.
\end{enumerate}
\begin{comment}
 \begin{enumerate}
\item \label{case:1} First consider the case $d \leq 0$. There are two further subcases.
 \begin{enumerate}
  \item \label{case:1a} Case $\alpha \not \perp \alpha'$. 
  Without loss of generality we may assume that $\alpha' = \alpha + \beta$
  for some $\rA_3$-triple $(\alpha, \beta, \gamma)$.
  Using~\eqref{eq:H1iii} and~\eqref{eq:bm} we obtain
   \begin{multline} \nonumber
   [S_\alpha(\xi), S_{\alpha+\beta}(bt^{m+1})] = [x_\alpha(\xi), [x_{\alpha+\beta+\gamma}(b't^m), x_{-\gamma}(t)]] = \\   
  = [[x_\alpha(\xi), x_{\alpha+\beta+\gamma}(b't^m)], {}^{x_{\alpha+\beta+\gamma}(b't^m)}\!x_{-\gamma}(t)] = 1 \text{ for some $b'\in A$ }.
  \end{multline}
  What this really proves: $PERP(d, 1)$ and $NONPERP(d, m)$ imply $NONPERP(d, m+1)$.
  \item Case $\alpha \perp \alpha'$. 
  As before, we set $\alpha' = \gamma$ and use~\eqref{eq:H1iii},\eqref{eq:bm} and~\eqref{eq:cm}:
  \begin{multline} \nonumber
   [S_\alpha(\xi), S_{\gamma}(bt^{m+1})] = [S_\alpha(\xi), [x_{\beta+\gamma}(b't^m), x_{-\beta}(t)]] = \\   
  = [[S_\alpha(\xi), x_{\beta+\gamma}(b't^m)], {}^{x_{\beta+\gamma}(b't^m)}\!x_{-\beta}(t)] = {}^{x_{\beta+\gamma}(b't^m)}\![x_{\alpha+\beta+\gamma}(b't^m\xi), x_{-\beta}(t)] = 1.
  \end{multline}
 \end{enumerate}
